---
title: "Social and Cultural Dynamics Exam ABM"
author: "Rebecca and Katrine"
output: 
  md_document:
    variant: markdown_github
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#INTRODUCTION 
This script computes an Agent-Based Model in which a group of agents has to solve a task, either in groups where they have the same level of ability of mixed levels. 

The model needs 4 fixed parameter: 
- N_agents = number of agents 
- rounds = number of rounds 
- group_condition = Grouped by mixed levels or similar levels # not there yet

When you run the model it outputs a dataframe containing the score for each agent on every round. 
From this dataframe you can plot how the agents performed given the fixed parameters. 

The model uses 4 predefined functions: 
- simulation() 
- getScore()
- runTask()
Each of these functions are defined below in the script along with explanations of what they do.
# Setup
```{r Setup}
library(pacman)
p_load(ggplot2, tidyverse) #Momocs - package that might not need to be used
#p_load(gganimate)
#library(gifski)
library(reshape2)
#library(crunch)

```


# Making groups
```{r Group-making}
# We want to define the error rate
#error_string <- round(c(runif(24, 0.1, 0.75)), 2)
#df <- as.data.frame(round(c(runif(24, 0.1, 0.75)), 2))

# We define the error rate as four different ones. 
group_n <- c(1, 2, 3, 4)
newbie <- round(c(runif(4, 0.45, 0.55)), 2)
novice <- round(c(runif(4, 0.55, 0.65)), 2)
master <- round(c(runif(4, 0.65, 0.75)), 2)
expert <- round(c(runif(4, 0.75, 0.85)), 2)

df_mixed <- data.frame(group_n, newbie, novice, master, expert)

# Now group_n is the mixed groups
# We define a group_n for the similar groups with the numbers from 5-8
group_n <- c(5, 6, 7, 8)

# We make a df for same level groups
df_similar <- as.data.frame(rbind(newbie, novice, master, expert))
df_similar$group_n <- group_n
df_similar <- melt(df_similar, id.vars = c("group_n"))
df_similar$id <- seq.int(nrow(df_similar))
# We make the variable column
variable <- rep(c("newbie", "novice", "master", "expert"), 4)
df_similar$variable <- variable

# Now the similar df is done

# Specify id.vars: the variables to keep but not split apart on
df_mixed <- melt(df_mixed, id.vars=c("group_n"))
df_mixed$id <- seq.int(from = 17, to = 32)
#reorder by column index
df_mixed <- df_mixed[c(4,1,3,2)]


```


#SIMULATION FUNCTION - simulation()

How the function works: 
- The simulation function takes the four inputs described above. 

- Generate a dataframe containing 8 columns: 
1) id, 2) group_n 3) variable 4) value 5) round 6) score of the agent. 

- Generate a matrix storing the guesses (combination of digits) of each agent in each round. The matrix has 3 dimensions (agent, digits, round). That is, for each round we have a 2D matrix containing N rows (one for each agent) and 15 columns (one for each digit). To index the guess of agent 3 on round 5: guesses[3, ,5].  
    
- Run loop to set op the values of the first round. For each agent, the loop does three things: 1) give the agent a first guess, by randomly sampling 15 digits between 0 and 9, 2) get the score of that guess (using the getScore() function) and add the score to the df, and 3) add the guess of the each agent to the guesses list of the first round. 

- Run learning. For each round, the loop does 3 things:  1) run task and update the df and guesses matrix.

- Return the dataframe (df) after running all rounds. 

```{r}
#df, guesses, og værdier for første runde
rounds <- 30
N_agents <- 32
N_groups <- 8

simulation <- function(N_agents,rounds, teaching){
  
  #Create df for round number, score etc
  df <-  data.frame(id = rep(1:N_agents, rounds), round = rep(1:rounds, each = N_agents), score = NA, Teaching_or_learning = NA, teaching_effect = teaching)
  df_groups <- rbind(df_similar, df_mixed)
  df <- cbind(df, df_groups)
  df <- df[c(1,6, 7,8, 2,3,4, 5)]
  
  # 3-Dimensional array to keep track of the agent's guesses - 
  # Dimensions [agent, digits, round]
  guesses = array(0,c(N_agents,50,rounds))   

  #Setup values for first round 
  for(agent in 1:N_agents){
    firstGuess <- round(runif(50, 0, 9) ,0) 
    guesses[agent,,1] <- firstGuess
    df$score[df$round==1 & df$id == agent] <- getScore(firstGuess)
  }
 ### Set up other columns in df here
  # We want a column specifying the group type
  df$group_type <- ifelse(df$group_n < 5, "Heterogeneous", "Homogeneous")
  # We want a mean for each group type 
  df$mean_group[df$round==1 & df$group_n < 5] <- mean(df$score[df$round==1 & df$group_n<5])
  df$mean_group[df$round==1 & df$group_n > 4] <- mean(df$score[df$round==1 & df$group_n>4])
  # We want a mean for each variable
   df$mean_variable[df$round==1 & df$variable == "newbie" & df$group_type ==  "Homogeneous"] <- mean(df$score[df$round==1 & df$variable == "newbie"  & df$group_type ==  "Homogeneous"])
   
  df$mean_variable[df$round==1 & df$variable == "novice" & df$group_type ==  "Homogeneous"] <- mean(df$score[df$round==1 & df$variable == "novice" & df$group_type ==  "Homogeneous"])
  
  df$mean_variable[df$round==1 & df$variable == "master" & df$group_type ==  "Homogeneous"] <- mean(df$score[df$round==1 & df$variable == "master" & df$group_type ==  "Homogeneous"])
  
  df$mean_variable[df$round==1 & df$variable == "expert" & df$group_type ==  "Homogeneous"] <- mean(df$score[df$round==1 & df$variable == "expert" & df$group_type ==  "Homogeneous"])
  
  df$mean_variable[df$round==1 & df$variable == "newbie" & df$group_type ==  "Heterogeneous"] <- mean(df$score[df$round==1 & df$variable == "newbie"  & df$group_type ==  "Heterogeneous"])
  
  df$mean_variable[df$round==1 & 
                     df$variable == "novice" & 
                     df$group_type ==  "Heterogeneous"] <- 
    mean(df$score[df$round==1 & df$variable == "novice" & df$group_type ==  "Heterogeneous"])
  
  df$mean_variable[df$round==1 & df$variable == "master" & df$group_type ==  "Heterogeneous"] <-mean(df$score[df$round==1 & df$variable == "master" & df$group_type ==  "Heterogeneous"])
  
  df$mean_variable[df$round==1 & df$variable == "expert" & df$group_type ==  "Heterogeneous"] <- mean(df$score[df$round==1 & df$variable == "expert" & df$group_type ==  "Heterogeneous"])
  
  #Run learning
  for(round in 2:rounds){

    task <- runTask(df, guesses, round, N_agents, teaching)
    df <- task$df
    guesses <- task$guesses

     # We want a mean for each group type 
  df$mean_group[df$round==round & df$group_n < 5] <- mean(df$score[df$round==round & df$group_n<5])
  df$mean_group[df$round==round & df$group_n > 4] <- mean(df$score[df$round==round & df$group_n>4])
  # We want a mean for each variable
  df$mean_variable[df$round==round & 
                     df$variable == "newbie" & df$group_type ==  "Homogeneous"] <- 
    mean(df$score[df$round==round & 
                    df$variable == "newbie" & df$group_type ==  "Homogeneous"])
  df$mean_variable[df$round==round 
                   & df$variable == "novice" & df$group_type ==  "Homogeneous"] <- 
    mean(df$score[df$round==round & 
                    df$variable == "novice" & df$group_type ==  "Homogeneous"])
  df$mean_variable[df$round==round & 
                     df$variable == "master" & df$group_type ==  "Homogeneous"] <- 
    mean(df$score[df$round==round & 
                    df$variable == "master" & df$group_type ==  "Homogeneous"])
  df$mean_variable[df$round==round & 
                     df$variable == "expert" & df$group_type ==  "Homogeneous"] <- 
    mean(df$score[df$round==round & 
                    df$variable == "expert" & df$group_type ==  "Homogeneous"])
   df$mean_variable[df$round==round & 
                      df$variable == "newbie" & df$group_type ==  "Heterogeneous"] <- 
    mean(df$score[df$round==round & 
                    df$variable == "newbie"  & df$group_type ==  "Heterogeneous"])
  df$mean_variable[df$round==round & df$variable == "novice" & df$group_type ==  "Heterogeneous"] <- 
    mean(df$score[df$round==round & df$variable == "novice" & df$group_type ==  "Heterogeneous"])
  df$mean_variable[df$round==round & df$variable == "master" & df$group_type ==  "Heterogeneous"] <- 
    mean(df$score[df$round==round & df$variable == "master" & df$group_type ==  "Heterogeneous"])
  df$mean_variable[df$round==round & df$variable == "expert" & df$group_type ==  "Heterogeneous"] <- 
    mean(df$score[df$round==round & df$variable == "expert" & df$group_type ==  "Heterogeneous"])
  
     # We want a mean endscore for all hetero and all homo groups
  df$mean_endscore[df$group_n < 5 & df$round == round] <- 
    mean(df$score[df$group_n<5 & df$round == rounds])
  df$mean_endscore[df$group_n > 4 & df$round == round] <- 
    mean(df$score[df$group_n>4 & df$round == rounds])
  }
  
  return(df)
}

```

#Task FUNCTION - runTask() 

How the function works: 
- It takes the df, guesses, round and N_agents as input 
- Loop through all groups
- Loop through all the agents in each group
- Create a new df containing only the values from the previous round. 
- Extract the previous score of that agent
- Extract the previous guess from that agent 
- Find the best scoring agent
- If yourself, you gain a higher probability of matching the string
- If another you take his/her guess and change one incorrect number
- If all of you have the best guess, all change one incorrect number
- Compute the score of the new guess 
- Add the new score to the df 
- Add the new guess to the guesses matrix 
- Return the updated df and guesses matrix 


```{r}
round <- 27
group <-  8
i <- 1

i
# i= 4, group = 8, round = 24 
# Error in if (newGuess[k] != target_string[k]) { : 
  #missing value where TRUE/FALSE needed
group
round

#Function to run tasks
runTask <- function(df, guesses, round, N_agents, teaching){
# prevround used to be here
  for (group in 1:length(unique(df$group_n))){
    df_group <- subset(df, group_n == group)
    for (i in 1:length(unique(df_group$id))){ # 
      prevRound <- df_group[df_group$round==round-1,]
      prevScore <- prevRound$score[i] 
      prevGuess <- guesses[prevRound$id[i],,round-1]
      # Draw random number between 0-9 for each group member only for i = 1
      if(i == 1) {
          draw_uncertainty <- c(sample(0:9,4, replace=TRUE))
          # Save teaching/learning as no if the number is 9 for each agent 
          prevRound$Teaching_or_learning[i] <- ifelse(draw_uncertainty[1] == 9, 
                                                      "No", "Yes")
          prevRound$Teaching_or_learning[i+1] <- ifelse(draw_uncertainty[2] == 9, 
                                                      "No", "Yes")
          prevRound$Teaching_or_learning[i+2] <- ifelse(draw_uncertainty[3] == 9, 
                                                      "No", "Yes")
          prevRound$Teaching_or_learning[i+3] <- ifelse(draw_uncertainty[4] == 9, 
                                                      "No", "Yes")
          # Save it to the overall df
          df$Teaching_or_learning[df$round == round-1 & 
                 df$id == prevRound$id[i] & 
                 df$group_n == group] <- prevRound$Teaching_or_learning[i]
          
          df$Teaching_or_learning[df$round == round-1 & 
                 df$id == prevRound$id[i+1] & 
                 df$group_n == group] <- prevRound$Teaching_or_learning[i+1]
          
          df$Teaching_or_learning[df$round == round-1 & 
                 df$id == prevRound$id[i+2] & 
                 df$group_n == group] <- prevRound$Teaching_or_learning[i+2]
          df$Teaching_or_learning[df$round == round-1 & 
                 df$id == prevRound$id[i+3] & 
                 df$group_n == group] <- prevRound$Teaching_or_learning[i+3]
          
          #Save it to the df_group
          df_group$Teaching_or_learning[df_group$round == round-1 & 
                 df_group$id == prevRound$id[i]] <- prevRound$Teaching_or_learning[i]
          
          df_group$Teaching_or_learning[df_group$round == round-1 & 
                 df_group$id == prevRound$id[i+1]] <- prevRound$Teaching_or_learning[i+1]
          
          df_group$Teaching_or_learning[df_group$round == round-1 & 
                 df_group$id == prevRound$id[i+2]] <- prevRound$Teaching_or_learning[i+2]
          df_group$Teaching_or_learning[df_group$round == round-1 & 
                 df_group$id == prevRound$id[i+3]] <- prevRound$Teaching_or_learning[i+3]
            
      }
        # check whether you should cooperate or work individually
      # Maybe change this to use prevround
      if (prevRound$Teaching_or_learning[i] == "No"){ # 10% of times agents do individual learning
            # Use your own previous guess
            newGuess <- prevGuess 
           
            # Change numbers that is not matching with target string
            stop <- 0
            k <- 1
            while (stop == 0){
              if (newGuess[k] != target_string[k]){
                Diff <- newGuess-target_string
                prevGuess_k <- round(newGuess[k]-(prevRound$value[i]/1)*Diff[k],0) 
                # They slowly get closer to the target value for each round. 
                newGuess[k] <- prevGuess_k
                stop <- 1
                } # changes a number in the string --> stops --> next round 
              else{ # If guess[k] = target[k] then it takes k + 1
                k = k+1 # Next number in the guess
              }
              }
          }
      else{ # The other 90% of the time they will cooperate

          #The code finds the guess of the best members (highest score) in previous round
            # We need to use the highest score of member you have a "yes" in teaching or learning
        # A max score df for everyone cooperating or not to use for the next if statement
            all_score <- subset(prevRound, prevRound$score == max(prevRound$score))
            maxScore_df <- as.data.frame(
              subset(prevRound, Teaching_or_learning == "Yes"))
            maxScore_df <- subset(maxScore_df, maxScore_df$score == max(maxScore_df$score))
            #maxScore_df <- subset(maxScore_df, Teaching_or_learning == "Yes")
          # Takes out one row of max scores in case of multiple
            maxScore <- maxScore_df[sample(nrow(maxScore_df), 1), ]
   
          # if you have the highest score last time and if not all members have it and you are a heterogenous group
            
            if(prevRound$group_n[i] <= 4 &
               nrow(all_score) < 4 & 
               maxScore$score == prevRound$score[i]){
              newGuess <- prevGuess 
          # Means you wait till everyone has the same as you. 
            # Can be changes later if we want to
             ## The effect of teaching
          # We will overwrite the value (error rate) for this agent like +0.01
            # Maybe not till next round ?? Right now it's from the current round
              for (error_round in round:rounds) {
                df$value[df$id == prevRound$id[i] &
                           df$round == error_round &
                           df$group_n == group] <- prevRound$value[i] + teaching
            }
          }
    
          #If you all score highest of if you did and not all did
            else{
          #Use the guess of the best scoring agent 
              newGuess <- guesses[maxScore$id,,round-1] #Index the guess of the best agent
            # Change numbers that is not matching with target string
                          
                          
              stop <- 0
              k <- 1
              while (stop == 0){
                if (newGuess[k] != target_string[k]){
                  Diff <- newGuess-target_string
                  prevGuess_k <- round(newGuess[k]-(prevRound$value[i]/1)*Diff[k],0) 
                 # They slowly get closer to the target value for each round. 
                  newGuess[k] <- prevGuess_k
                  stop <- 1
                  } # changes a number in the string --> stops --> next round 
                else{ # If guess[k] = target[k] then it takes k + 1
                  k = k+1 # Next number in the guess
                  #if(k==15){ #It has now guessed the string and therefore stops
                    #stop <- 1
                #}
              }
              }
          }
          }
          # We save a score for the new guess
          score <- getScore(newGuess)
          #We save the newguess into the guesses matrix 
          guesses[prevRound$id[i],,round] <- newGuess # newGuess into the guesses matrix
          # Save it to the overall df
          df$score[df$round == round & 
                 df$id == prevRound$id[i] & 
                 df$group_n == group] <- score
        }
    }
  output <- list(df=df, guesses=guesses)
  return(output)
}

```


#GET SCORE FUNCTION - getScore()

How the function works: 
- It takes a guess (list of 15 digits) as input
- Set the score to 0 
- Loop through each of the 15 digits: 
  - If the digit is equal to the same digit in the target string, then add 1 to the score. 
  - If not, do nothing, and continue with the next digit. 
- Return the score. 

```{r Get score}
#Function to calculate score 
getScore <- function(guess){
  score = 0
  for(i in 1:50){
    if(guess[i] == target_string[i]){
      score = score+1
      i = i+1
    }
  }
  return(score)
}
```


#RUN SIMULATION 

- First, we setup the task. The task is defined as a string of 15 digtis between 0 and 9. The agents have to figure out the correct combination of digits. On every round, they produce one guess.  

- Second, we run a simulation using the simulation() function. The simulation function takes 4 inputs: N_agents = number of agents, rounds = number of rounds. The function outputs a dataframe, containing the performance of each agent on each round.

```{r Run simulation}
#Setup task 
target_string = c(sample(0:9,50, replace=TRUE))

# Run simulation
Class_005 <- simulation(
  N_agents <- 32,
  rounds <- 40, # Can be changed
  teaching <- 0.005 # Can be changed 
)

write.csv(Class_005, "Class_005_2.csv")

# Run simulation
Class_006 <- simulation(
  N_agents <- 32,
  rounds <- 40, # Can be changed
  teaching <- 0.006 # Can be changed 
)

write.csv(Class_006, "Class_006_2.csv")

# Run simulation
Class_007 <- simulation(
  N_agents <- 32,
  rounds <- 40, # Can be changed
  teaching <- 0.007 # Can be changed 
)

write.csv(Class_007, "Class_007_2.csv")

# Run simulation
Class_008 <- simulation(
  N_agents <- 32,
  rounds <- 40, # Can be changed
  teaching <- 0.008 # Can be changed 
)

write.csv(Class_008, "Class_008_2.csv")

# Run simulation
Class_009 <- simulation(
  N_agents <- 32,
  rounds <- 40, # Can be changed
  teaching <- 0.009 # Can be changed 
)

write.csv(Class_009, "Class_009_2.csv")

# Run simulation
Class_01 <- simulation(
  N_agents <- 32,
  rounds <- 40, # Can be changed
  teaching <- 0.01 # Can be changed 
)

write.csv(Class_01, "Class_01_2.csv")

# Run simulation
Class_02 <- simulation(
  N_agents <- 32,
  rounds <- 40, # Can be changed
  teaching <- 0.02 # Can be changed 
)

write.csv(Class_02, "Class_02_2.csv")

# Run simulation
Class_03 <- simulation(
  N_agents <- 32,
  rounds <- 40, # Can be changed
  teaching <- 0.03 # Can be changed 
)

write.csv(Class_03, "Class_03_2.csv")

# Run simulation
Class_04 <- simulation(
  N_agents <- 32,
  rounds <- 40, # Can be changed
  teaching <- 0.04 # Can be changed 
)

write.csv(Class_04, "Class_04_2.csv")

# Run simulation
Class_05 <- simulation(
  N_agents <- 32,
  rounds <- 40, # Can be changed
  teaching <- 0.05 # Can be changed 
)

write.csv(Class_05, "Class_05_2.csv")



```

#PLOT RESULTS 
```{r}

#Combine Best Member data 
bmData <- rbind(Class_005, Class_006, Class_007, Class_008, Class_009, Class_01, Class_02, Class_03, Class_04, Class_05) 
bmData

bmData$teaching_effect <- as.factor(bmData$teaching_effect)

#write.csv(bmData, "bmData1.csv")


##################################################################################################
#PLOTS 

# We want to plot overall score per each round seperately for homo and hetero groups
ggplot(Class_01, aes(x = round, y = mean_group, color = group_type)) + 
  geom_point() + geom_smooth(method = "lm") +
  labs(title = "Mean score of the hetero and homo groups per round")


# We want to plot the overall score for experts per group
ggplot(Class_01[Class_01$variable == "expert",], aes(x = round, y = mean_variable, color = group_type)) + geom_point() + geom_smooth(method = "lm") + labs(title = " Mean score of the experts per round in the two group types")

# We want to plot the overall score for newbies per group
ggplot(Class_005[Class_005$variable == "newbie",], aes(round, score, color = group_type)) + geom_point() + geom_smooth(method = "lm") + labs(title = " Mean score of the newbies per round in the two group types")
View(Class_005)

# We want to plot the overall score for novice per group
ggplot(Class_01[Class_01$variable == "novice",], aes(x = round, y = mean_variable, color = group_type)) + geom_point() + geom_smooth(method = "lm") + labs(title = " Mean score of the experts per round in the two group types")

# We want to plot the overall score for master per group
ggplot(Class_005[Class_005$variable == "master",], aes(round, score, color = group_type)) + geom_point() + geom_smooth(method = "lm") + labs(title = " Mean score of the newbies per round in the two group types")
View(Class_005)

# We want to plot the overall score for each variable per group
ggplot(Class_01, aes(round, mean_variable, color = group_type)) + geom_point() + geom_smooth(method = "lm") + facet_wrap(~variable) 

# We want to see how the effect of teaching changes the endscore
ggplot(bmData, aes(x = teaching_effect, y = mean_endscore, fill = group_type)) + geom_bar(stat = "identity", position = "dodge") 


## Effect of teaching per round
ggplot(Class_005, aes(x = round, y = value, color = variable)) + geom_point()

##################################################################################################
#PLOT with gganimate - GIF
#plotLearningStrategyProp20 <- ggplot(allData[allData$prop_ind==0.2,], aes(round, meanAll))+ 
  # geom_line(aes(color=SL_strategy)) + 
  # labs(title = "Mean score of the group per round")+
  # transition_reveal(round) #gganimate, reveal results gradually as a function of round


#Create gif of the animation 
#anim_plotLearningStrategyProp20 <-animate(plotLearningStrategyProp20,100,fps = 20,duration = 30, width = 950, height = 750, renderer = gifski_renderer(loop = FALSE))

#Save the gif 
#anim_save("anim_plotLearningStrategyProp20.gif", animation=anim_plotLearningStrategyProp20)

#anim_plotLearningStrategyProp20


round(4-(0.45/2)*-4,0)
round(5-(0.45/2)*-1,0)
round(5-(0.45/2)*-1,0)

round(4-(0.55/2)*-4,0)
round(5-(0.55/2)*-1,0)
round(5-(0.65/2)*-1,0)
round(5-(0.65/2)*-1,0)

round(4-(0.65/1)*-4,0)
round(7-(0.65/1)*-1,0)


round(4-(0.75/1)*-4,0)
round(7-(0.75/1)*-1,0)


round(4-(0.85/1)*-4,0)
round(7-(0.85/1)*-1,0)


round(4-(0.80)*-4,0)
round(7-(1)*-1,0)



```





